Towards a reproducible framework for undergraduate data science research

### 2019-11-29 Rmd/introduction.Rmd
udpated intro & README

@nolan2010computing fervently argue for students to work with real data. 
Working with real data allows them to develop skill not only in statistical analysis,
but also in data transfer from online sources, in data storage, and in using data 
from multiple file formats. In the case of Twitter data, tweets are stored in
Javascript Object Notation (JSON) [@tweet_json, @json].

A key component of our mentoring framework is the emphasis on using real data to answer real 
scientific questions. We believe that this process develops the problem-solving skills that 
students will need in their future careers in data science.

We encouraged the student to articulate a scientific research question, translate that question into quantitative and statistical terms, determine which data could be used to address the question, acquire the data, analyze data, visualize results, and communicate what they learned.

@@

>>>>

@nolan2010computing argue for students to work with real data. 
Working with real data allows them to develop skill not only in statistical analysis,
but also in data transfer from online sources, in data storage, and in using data 
from multiple file formats. In the case of Twitter data, tweets are stored in
Javascript Object Notation (JSON) [@tweet_json, @json].

Mentoring in the work place and in higher education can have many benefits, including [@higgins2001reconceptualizing].
A key component of our mentoring framework is the emphasis on using real data to answer real 
scientific questions. We believe that this process develops the problem-solving skills that 
students will need in their future careers in data science. 
We encouraged the student to articulate a scientific research question, translate that question 
into quantitative and statistical terms, determine which data could be used to address the 
question, acquire the data, analyze data, visualize results, and communicate what they learned.

We provide guidance regarding selection of

@@

## Our backgrounds

During the time when we first implemented our framework, we served as early-career instructors in 
the statistics department at the University of Wisconsin-Madison. One of us (Hanlon) had prior 
experience in mentoring undergraduates, while the other (Boehm) had none. Our initial 
conceptualization of mentoring drew heavily on ideas we first encountered in professional development 
courses, including the Delta Program's Entering Mentoring class (https://delta.wisc.edu). 
Professor Erik Nordheim heavily 
influenced our approach to and philosophy of teaching statistics. 
We studied with Professor Nordheim early in our teaching 
careers, and his guidance continues to influence our teaching practices.

We both have experience in teaching undergraduate introductory statistics courses with enrollments over 
100 students. Through our interactions with students in these classes, we've come to value not only 
the ideas in a traditional introductory course, but also the need to prepare students with the essential skills needed for success in data science. @nolan2010computing summarizes these skill sets in the following three ideas:

1. broaden statistical computing to include emerging areas
1. deepen computational reasoning skills
1. combine computational topics with data analysis in the practice of statistics

To these three praiseworthy goals, we add a fourth:

4. develop skills in reproducible research to promote the practice of open science

We see the fourth goal as an equal with the first three from @nolan2010computing. 
Data scientists are uniquely positioned to promote open science practices, including the free sharing of 
data, code, and instructions for their use. The recently recognized need for science to be more transparent and more reproducible elevate this goal to the level of the first three.

Below, we detail our methods for creating a reproducible framework for undergraduate data science 
research. We describe our results before concluding with lessons learned, 
things we could have done differently, and recommendations for future 
mentors who may use and extend our framework.

### 2019-11-29 Rmd/methods.Rmd
udpated intro & README

Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research. Given the growing imperative to document and share code, we feel that this skill set equals in importance the three points above.

4. develop and practice skills in reproducible research

Below, we describe how our framework enabled students to achieve in the four items listed above.

@@

data analysis and practice of statistics:

>>>>

Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research. Given the growing imperative to document and share code to promote open science, we feel that this skill set equals in importance the three points above.

4. develop and practice skills in reproducible research to promote open science

Below, we describe how our framework enabled students to achieve competence in the four areas listed above.

@@

### data analysis and practice of statistics
